Constant-round Blind Classical Verification of Quantum Sampling

### 2020-08-01 preliminaries.tex
Reorganize prelim

Let $\mathcal{B}$ be the Hilbert space corresponding to one qubit. Let $H:\mathcal{B}^{\otimes n}\rightarrow\mathcal{B}^{\otimes n}$ be Hermitian matrices. We use $H\geq0$ to denote $H$ being positive semidefinite. Let $\lambda(H)$ be the smallest eigenvalue of $H$. The ground states of $H$ are the eigenvectors corresponding to $\lambda(H)$. For matrix $H$ and subspace $S$, let $H\big|_S=\Pi_S H \Pi_S$, where $\Pi_S$ is the projector onto the subspace $S$.
%For a $T$-qubit Hilbert space, let the state $\ket{\widehat{t}}=\ket{1}^{\otimes t}\otimes \ket{0}^{{\otimes (T-t)}}$.

For $n$-qubit states $\rho_1, \rho_2\in\cB^{\otimes n}$,
let $\norm{\rho_1-\rho_2}_{tr}=\frac{1}{2}\norm{\rho_1-\rho_2}_1$ denote the trace distance between $\rho_1$ and $\rho_2$. We say $\rho_1$ and $\rho_2$ are $\eps$-close if $\norm{\rho_1- \rho_2}_{tr}\leq\eps$.
For two distributions $f_1$ and $f_2$ over the same finite domain $X$, let $\norm{f_1-f_2}_{TV}=\frac{1}{2}\sum_{x\in X}\abs{f_1(x)-f_2(x)}$ denote their total variation distance,
and we say $f_1$ and $f_2$ are $\eps$-close if $\norm{f_1-f_2}_{TV}\leq\eps$.
We write $x\xleftarrow{\$}X$ when $x$ is sampled uniformly at random from the set $X$.

\begin{definition} [quantum-classical channels]
    \label{def:QCChannel}
    A quantum measurement is given by a set of matrices $\set{M_k}$ such that $M_k\geq0$ and $\sum_k M_k=\id$.
    We associate to any measurement a map $\Lambda(\rho)=\sum_k \tr(M_k\rho)\proj{k}$
    with $\{\ket{k}\}$ an orthonormal basis.
    This map is also called a \emph{quantum-classical channel}.
\end{definition}

@@

\Ethan{Quantum measurements...}

\begin{definition}[$M_{XZ}(\rho,h)$]
	For any natural number $n$, $n$-bit string $h$, and $n$-qubit quantum state $\rho$, consider the following measurement procedure: measure the first qubit of $\rho$ in $X$ basis if $h_1=0$; measure the first qubit of $\rho$ in $Z$ basis if $h_1=1$.  Measure the second qubit of $\rho$ in $X$ basis if $h_2=0$; measure the second qubit of $\rho$ in $Z$ basis if $h_2=1$. Continue qubit-by-qubit until all $n$-qubits of $\rho$ are measured, where $i$-th qubit is measured in $X$ basis if $h_i=0$ and  $i$-th qubit is measured in the $Z$ basis if $h_i=1$.

@@

\begin{definition}
    \label{dfn:stats-secure-proto-sampbqp}
    Let $\Pi=(P, V)$ be a $\QPIP_\tau$ protocol.
    We say it is a protocol for the $\SampBQP$ instance $(D_x)_{x\in\zo^*}$ with completeness error $c(\cdot)$ and statistical (resp., computational) soundness if the following holds:
    \begin{itemize}
        \item On public inputs $1^\lambda$, $1^{1/\eps}$, and $x\in\zo^{\poly(\lambda)}$, $V$ outputs $(d, z)$ where $d\in\set{\Acc, \Rej}$.
            If $d=\Acc$ then $z\in\zo^{m(\abs{x})}$ where $m$ is given in \Cref{dfn:sampling-problem}, otherwise $z=\bot$.
        \item (Completeness):
            For all accuracy parameter $\eps(\lambda)=\frac{1}{\poly(\lambda)}$,
            security parameter $\lambda\in\bbN$,
            and $x\in\zo^{\poly(\lambda)}$,
            Let $(d, z)\leftarrow(P, V)(1^\lambda, 1^{1/\eps}, x)$, then $d=\Rej$ with probability at most $c(\lambda)$.
        \item (Statistical soundness): For all cheating prover $P^*$,
            accuracy parameter $\eps(\lambda)=\frac{1}{\poly(\lambda)}$,
            sufficiently large $\lambda\in\bbN$, and $x\in\zo^{\poly(\lambda)}$,
            consider the following experiment:
            \begin{itemize}
                \item Let $(d, z)\leftarrow(P^*, V)(1^\lambda, 1^{1/\eps}, x)$.
                \item Define $z_{ideal}$ by
                $$\begin{cases}
                    z_{ideal}=\bot & \text{if } d=\Rej\\
                    z_{ideal}\leftarrow D_x & \text{if } d=\Acc
                \end{cases}$$.
            \end{itemize}
            It holds that $\norm{(d,z)-(d,z_{ideal})}_{TV}\leq\eps$.
        \item (Computational soundness):
        For all cheating $\BQP$ prover $P^*$, $\BQP$ distinguisher $\mathsf{D}$, accuracy parameter $\eps(\lambda)=\frac{1}{\poly(\lambda)}$,
            sufficiently large $\lambda\in\bbN$, and all $x\in\zo^{\poly(\lambda)}$,
            define $d, z, z_{ideal}$ by the same experiment as above.
            It holds that
            $$\abs{\Pr[\mathsf{D}(1^{\lambda}, d,z) = 1] - \Pr[\mathsf{D}(1^{\lambda}, d,z_{ideal}) = 1]} \leq \eps $$
    \end{itemize}
\end{definition}

@@

We now define blindness of a protocol.
In the context of blind protocols, we consider $x$ as the verifier's private input instead of a common input.  
%the above definitions should be modified so that $x$ is the verifier's private input.
%We define blindness with an accuracy parameter $\eps$, which exists in protocols for $\SampBQP$ instances;
%one can leave it out where appropriate, such as in protocols for $\BQP$ languages.

\Ethan{The following is more crypto based

Others don't have a fully general definition,
or it's just statistical/OTP.

Need explanation/discussion for this.
Nobody else writes it like this,
but this is the correct way to write it.

Explain what is this saying.}

@@

\Ethan{Maybe swap 2.9 and 2.8 or move it up even more.
If it's before 2.7 then we can just write 2.7 in terms of this.

Consider making this its own section.}

\begin{definition}
\label{def:comp-indis}
    %\Ethan{TODO PROOFREAD ME}
    Two ensemble of states $\set{\rho_{1,\lambda}}_{\lambda\in\bbN}$ and $\set{\rho_{2,\lambda}}_{\lambda\in\bbN}$ are \emph{$\eps$-computationally indistinguishable} if for all $\BQP$ distinguishers $\mathsf{D}$ and $\lambda\in\bbN$,
    $$\abs{\Pr\left[\mathsf{D}(\rho_{1,\lambda})=1\right]-\Pr\left[\mathsf{D}(\rho_{2,\lambda})=1\right]}\leq\eps(\lambda).$$
    We also denote $\eps$-computational indistinguishability by $\rho_{1,\lambda}\approx_{c, \eps}\rho_{2,\lambda}$.
    We drop the $\eps$ parameter when it is negligible.
    We extend computational indistinguishability to subnormalized states by taking the distinguisher $\mathsf{D}$ to output subnormalized states as appropriate.
    We also extend it to distributions by treating them as mixed states.
\end{definition}

@@

\Ethan{Chernoff bound looks like it's in the middle of nowhere. Just get rid of it.}

\iffalse
\subsection{Chernoff bound}

\begin{thm}
\label{thm:Chernoff}
Let $X=\sum_{i=1}^n X_i$ where $X_i$ are independent Bernoulli trials, and $\mu=\E[X]$.
Then for all $0<\delta<1$,
$$P[\abs{X-\mu}\geq\delta\mu]\leq2e^{-\frac{\mu\delta^2}{3}}.$$
\end{thm}

>>>>

Let $\mathcal{B}$ be the Hilbert space corresponding to one qubit. Let $H:\mathcal{B}^{\otimes n}\rightarrow\mathcal{B}^{\otimes n}$ be Hermitian matrices. We use $H\geq0$ to denote $H$ being positive semidefinite. Let $\lambda(H)$ be the smallest eigenvalue of $H$. The ground states of $H$ are the eigenvectors corresponding to $\lambda(H)$. For matrix $H$ and subspace $S$, let $H\big|_S=\Pi_S H \Pi_S$, where $\Pi_S$ is the projector onto the subspace $S$.
%For a $T$-qubit Hilbert space, let the state $\ket{\widehat{t}}=\ket{1}^{\otimes t}\otimes \ket{0}^{{\otimes (T-t)}}$.
We write $x\xleftarrow{\$}X$ when $x$ is sampled uniformly at random from the set $X$.

@@

For $n$-qubit states $\rho_1, \rho_2\in\cB^{\otimes n}$,
let $\norm{\rho_1-\rho_2}_{tr}=\frac{1}{2}\norm{\rho_1-\rho_2}_1$ denote the trace distance between $\rho_1$ and $\rho_2$. We say $\rho_1$ and $\rho_2$ are $\eps$-close if $\norm{\rho_1- \rho_2}_{tr}\leq\eps$.
For two distributions $f_1$ and $f_2$ over the same finite domain $X$, let $\norm{f_1-f_2}_{TV}=\frac{1}{2}\sum_{x\in X}\abs{f_1(x)-f_2(x)}$ denote their total variation distance,
and we say $f_1$ and $f_2$ are $\eps$-close if $\norm{f_1-f_2}_{TV}\leq\eps$.
We additionally denote the following for states that are only computationally indistinguishable (but might be statistically far):
\begin{definition}
\label{def:comp-indis}
    Two ensemble of states $\set{\rho_{1,\lambda}}_{\lambda\in\bbN}$ and $\set{\rho_{2,\lambda}}_{\lambda\in\bbN}$ are \emph{$\eps$-computationally indistinguishable} if for all $\BQP$ distinguishers $\mathsf{D}$ and $\lambda\in\bbN$,
    $$\abs{\Pr\left[\mathsf{D}(\rho_{1,\lambda})=1\right]-\Pr\left[\mathsf{D}(\rho_{2,\lambda})=1\right]}\leq\eps(\lambda).$$
    We also denote $\eps$-computational indistinguishability by $\rho_{1,\lambda}\approx_{c, \eps}\rho_{2,\lambda}$.
    We drop the $\eps$ parameter when it is negligible.
    We extend computational indistinguishability to subnormalized states by taking the distinguisher $\mathsf{D}$ to output subnormalized states as appropriate.
    We also extend it to distributions by treating them as mixed states.
\end{definition}

We denote measurement as follows:
\begin{definition} [quantum-classical channels]
    \label{def:QCChannel}
    A quantum measurement is given by a set of matrices $\set{M_k}$ such that $M_k\geq0$ and $\sum_k M_k=\id$.
    We associate to any measurement a map $\Lambda(\rho)=\sum_k \tr(M_k\rho)\proj{k}$
    with $\{\ket{k}\}$ an orthonormal basis.
    This map is also called a \emph{quantum-classical channel}.
\end{definition}
Furthermore, we associate measurement outcomes in $X$ or $Z$ basis with a corresponding random variable as follows:
\begin{definition}[$M_{XZ}(\rho,h)$]
	For any natural number $n$, $n$-bit string $h$, and $n$-qubit quantum state $\rho$, consider the following measurement procedure: measure the first qubit of $\rho$ in $X$ basis if $h_1=0$; measure the first qubit of $\rho$ in $Z$ basis if $h_1=1$.  Measure the second qubit of $\rho$ in $X$ basis if $h_2=0$; measure the second qubit of $\rho$ in $Z$ basis if $h_2=1$. Continue qubit-by-qubit until all $n$-qubits of $\rho$ are measured, where $i$-th qubit is measured in $X$ basis if $h_i=0$ and  $i$-th qubit is measured in the $Z$ basis if $h_i=1$.

@@

\begin{definition}
    \label{dfn:stats-secure-proto-sampbqp}
    Let $\Pi=(P, V)$ be a $\QPIP_\tau$ protocol.
    We say it is a protocol for the $\SampBQP$ instance $(D_x)_{x\in\zo^*}$ with completeness error $c(\cdot)$ and statistical (resp., computational) soundness if the following holds:
    \begin{itemize}
        \item On public inputs $1^\lambda$, $1^{1/\eps}$, and $x\in\zo^{\poly(\lambda)}$, $V$ outputs $(d, z)$ where $d\in\set{\Acc, \Rej}$.
            If $d=\Acc$ then $z\in\zo^{m(\abs{x})}$ where $m$ is given in \Cref{dfn:sampling-problem}, otherwise $z=\bot$.
        \item (Completeness):
            For all accuracy parameter $\eps(\lambda)=\frac{1}{\poly(\lambda)}$,
            security parameter $\lambda\in\bbN$,
            and $x\in\zo^{\poly(\lambda)}$,
            Let $(d, z)\leftarrow(P, V)(1^\lambda, 1^{1/\eps}, x)$, then $d=\Rej$ with probability at most $c(\lambda)$.
        \item (Statistical soundness): For all cheating prover $P^*$,
            accuracy parameter $\eps(\lambda)=\frac{1}{\poly(\lambda)}$,
            sufficiently large $\lambda\in\bbN$, and $x\in\zo^{\poly(\lambda)}$,
            consider the following experiment:
            \begin{itemize}
                \item Let $(d, z)\leftarrow(P^*, V)(1^\lambda, 1^{1/\eps}, x)$.
                \item Define $z_{ideal}$ by
                $$\begin{cases}
                    z_{ideal}=\bot & \text{if } d=\Rej\\
                    z_{ideal}\leftarrow D_x & \text{if } d=\Acc
                \end{cases}$$.
            \end{itemize}
            It holds that $\norm{(d,z)-(d,z_{ideal})}_{TV}\leq\eps$.
		\item (Computational soundness): \Ethan{TODO write in terms of comp. indistinguishability?}
        For all cheating $\BQP$ prover $P^*$, $\BQP$ distinguisher $\mathsf{D}$, accuracy parameter $\eps(\lambda)=\frac{1}{\poly(\lambda)}$,
            sufficiently large $\lambda\in\bbN$, and all $x\in\zo^{\poly(\lambda)}$,
            define $d, z, z_{ideal}$ by the same experiment as above.
            It holds that
            $$\abs{\Pr[\mathsf{D}(1^{\lambda}, d,z) = 1] - \Pr[\mathsf{D}(1^{\lambda}, d,z_{ideal}) = 1]} \leq \eps $$
    \end{itemize}
\end{definition}

@@

We now define blindness of a protocol.
In the context of blind protocols, we consider $x$ as the verifier's private input instead of a common input.

\Ethan{The following blindness definition is more crypto-based, but since we're submitting it to Eurocrypt maybe this is fine as-is?}

@@

@@
